---
title: "Persona Discovery"
#author: "Ashish Singh (ashish@iashishsingh.com)"
date: "5/16/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
library(tidyverse)
library(lubridate)
library(readr)
library(ordinal)
library(cluster)

survey <- read_csv("~/Code/gaugedesign/data/viz/processed-dataset.csv")
```


## Visualize the variables
## subset data for use in the example

```{r}
survey_vars <- survey %>%
  select(
    `Employment Status`,
    Gender,
    Age,
    Education,
    HHI,
    `Own-Rent`,
    `Perception of Protection`
  )
```


## visualize the data
```{r}
survey_vars %>%
  gather(
    var,
    value
  ) %>%
  ggplot(
    aes(value, fill = var)
  ) +
  geom_histogram(stat = "count") +
  facet_wrap( var ~ ., scales = "free_x")
```


## Covert categorical fields to a factor datatype
```{r}
survey_vars$`Employment Status` <- as.factor(survey_vars$`Employment Status`)
survey_vars$Gender <- as.factor(survey_vars$Gender)
survey_vars$Age <- as.factor(survey_vars$Age)
survey_vars$Education <- as.factor(survey_vars$Education)
survey_vars$HHI <- as.factor(survey_vars$HHI)
survey_vars$`Own-Rent` <- as.factor(survey_vars$`Own-Rent`)
survey_vars$`Perception of Protection` <- as.factor(survey_vars$`Perception of Protection`)
```


## Pre-process the data and compute distance matrix

```{r}
survey_dist <- daisy(survey_vars)
```

## Find the optimal number of clusters

### hclust to see dentogram
```{r}
survey_hclust <- hclust(survey_dist)
plot(survey_hclust, labels = FALSE, main = "Default hclust")
```

### cut at the "top" of the tree

```{r}
group_3 <- cutree(survey_hclust, 3)
table(group_3)
```


### Using Elbow method to find no. of clusters

```{r}
# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(survey_dist, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k_values <- 1:15

# compute wss for 1-15 clusters. This can take some time.
wss_values <- map_dbl(k_values, wss)

plot(k_values, wss_values,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```


## Compute Clusters

```{r}
# from elbow evaluation choose cluster size (centers)
k5 <- kmeans(survey_dist, centers = 5, nstart = 25)

# view cluster distribution
table(k5$cluster)
## 
##   1   2   3   4   5 
## 179 192 219 195 103
# add the cluster classifications to the data observations.
survey$cluster <- factor(k5$cluster)

```

Five clusters produce reasonably balanced cluster sizes.

## Identify variables contributing to clusters

CLM is used for this example.

```{r}
# build model
fm_survey <- clm(cluster ~ `Employment Status` + Gender + Age + Education + HHI + `Own-Rent` + `Perception of Protection`, data = survey)
# examine significance of variables
summary(fm_survey)
```

## Review Pairwise Plots

```{r}
survey %>% 
  ggplot(
    aes(Age, Gender, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(`Employment Status`, Gender, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(`Employment Status`, Age, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(Education, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(`Employment Status`, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(Gender, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(Age, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(`Own-Rent`, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

```{r}
survey %>% 
  ggplot(
    aes(HHI, `Perception of Protection`, color = cluster)
  ) +
  geom_jitter()
```

## Summarize characteristics and define personas

```{r}
tribble(
  ~cluster, ~gender, ~perception_of_product, ~age_group, ~own_or_rent,
  #-------|--------|-----------------------|-----------
  "1", "male", "Makes sense for expensive", "25-55", "own",
  "2", "female", "Makes sense for expensive, Smart thing to do", "25-69", "rent",
  "3", "female", "All", "All", "own",
  "4", "male", "All", "18-54", "rent",
  "5", "female", "Makes sense for expensive", "25-59", "own"
)
```

